---
title: 'Homework #2'
author: "Jacob Shoemaker; jshoemaker2249"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework is about syntax and basic plotting functions in R, and is **due on Feb 5** before class.

**Submission instructions**: Create a folder named `ds202_hw2`, and name the RMarkDown file including your solutions `hw2.Rmd` under this folder. For submission, create a GitHub repository named `ds202_hw2` under your account, and push both `hw2.Rmd` and the knitted `hw2.html` to GitHub before the deadline. I will assume you use the same GitHub username as for your HW1 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. Some questions will require the following answer format.

## Example question and answer
0\. What is the smallest positive odd integer? Storage it in a variable named `a0` and print it.

```{r}
a0 <- 1
a0
```

## R commands
1. Today's temperature has a high of 15 and a low of -10 degree fahrenheit.  Find the corresponding degress in celsius and storage it in a (single) vector named `cel`. Print the variable.
```{r}
cel <- c(-9.44, -23.33)
cel
```
2. Create a vector of length 200. The odd entries should be 1 and even entries should be 0. Storage the created in a variable named `vec200` and print.
```{r}
vec200 <- rep(c(0,1), length.out = 200)
vec200
```
3. How many perfect squares are there between 1 and 1001? Storage the answer in a  variable named `numPerfect` and print. You may want to look at `?round`.
```{r}
numPerfect <- (1 + floor(sqrt(1001)) - ceiling(sqrt(1)))
numPerfect
```

## Cars and mileage

```{r}
library(ggplot2)
```
1. What are the 3 most fuel-efficient models in terms of highway mileage? Storage the data frame (with 3 rows) containing all information of these three cars in a variable named `top3` and print.

```{r}
top3 <- head(mpg,3)
top3
```

2. How many different compact models are included in this dataset? Storage the answer in  a variable named `numCompact` and print.
```{r}
numCompact <- nrow(mpg[mpg$class == 'compact', ])
numCompact
```
3. Create a scatterplot of `hwy` and `cty` and describe the relationship. Why are there so few points visible?
```{r}
qplot(hwy,cty,data=mpg)
```

###Why are there so few points visible?
Not all of the points are visible because there are some duplicate rows in which there are the same fuel efficiency stats.

4. Are cars in 2008 more fuel efficient than cars in 1999? Use graphs *and* summary statistics to answer this question. 

```{r}
#Breaking up the dataframe in two based on year
dat1999 <- mpg[mpg$year == '1999', ]
dat2008 <- mpg[mpg$year == '2008', ]
qplot(hwy,cty, data= dat1999)
qplot(hwy,cty, data= dat2008)
```


Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission. Make sure that you don't use any file structure that depends on your computing environment.